Type theory in type theory with single substitutions

### 2025-01-30 cpp/p.tex
stuff

\begin{definition}[Minimised single substitution calculus with $\Pi$ and $\U$]\label{def:min}
  This section defined the minimised first-order version of the SOGAT
  (\ref{eq:tt}), relying essentially on the presence of universes and
  $\Pi$ types. For reference, we list all the rules of the minimised
  calculus in Appendix \ref{app:minimisation}.
\end{definition}
\begin{problem}
  The GATs of Definition \ref{def:min} and Definition \ref{def:tt} are
  isomorphic, in particular, all equations are interderivable.
\end{problem}
\begin{proof}[Construction]
  Clear from the construction in this section.
\end{proof}

@@

    & \Con && := (i:\N)\times\Ty\,\diamond\,i \\
    & \Sub\,\Delta\,\Gamma && := \Tm\,\diamond\,(\Lift^{\Gamma-\Delta}\,\Delta\Ra\Lift^{\Delta-\Gamma}\,\Gamma)

@@

    & \gamma\circ\delta && := \lam\Bigg(\mk^{\Theta-\Gamma}\,\bigg(\un^{\Delta-\Gamma}\,\Big(\gamma[\p]\cdot\mk^{\Gamma-\Delta}\,\big(\un^{\Theta-\Delta}\, \\[-1.5em]
    & && \hspace{10em}(\delta[\p]\cdot\mk^{\Delta-\Theta}\,(\un^{\Gamma-\Theta}\,\q))\big)\Big)\bigg)\Bigg) \\
    & \rlap{$\id := \lam\,\q \hspace{4.35em} \diamond := (0,\top) \hspace{4.35em} \epsilon := \lam\,(\mk^\Gamma\,\tt)$}

@@

    & \Tm\,\Gamma\,A && := \Tm\,\diamond\,\Bigg(\Pi\,(\Lift^{i-\Gamma}\,\Gamma)\,\bigg(\Lift^{\Gamma-i}\,\Big(\El\, \\[-1em]
    & && \hspace{4em} \big(\un^{\Gamma-(1+i)}\,(A[\p]\cdot\mk^{1+i-\Gamma}\,(\un^{i-\Gamma}\,\q))\big)\Big)\bigg)\Bigg)

@@

    & \Con && := \Ty\,\diamond \\
    & \Sub\,\Delta\,\Gamma && := \Tm\,\diamond\,(\Delta\Ra\Gamma) \\
    & \gamma\circ\delta && := \lam\,(\gamma[\p]\cdot(\delta[\p]\cdot\q)) \\
    & \id && := \lam\,\q \\
    & \diamond && := \top \\
    & \epsilon && := \lam\,\tt \\
    & \Ty\,\Gamma && := \Tm\,\diamond\,(\Gamma\Ra\U) \\
    & A[\gamma] && := \lam\,(A[\p]\cdot(\gamma[\p]\cdot\q)) \\
    & \Tm\,\Gamma\,A && := \Tm\,\diamond\,(\Pi\,\Gamma\,(\El\,(A[\p]\cdot\q))) \\
    & a[\gamma] && := \lam\,(a[\p]\cdot(\gamma[\p]\cdot\q)) \\
    & \Gamma\ext A && := \Sigma\,\Gamma\,(\El\,(A[\p]\cdot\q)) \\
    & (\gamma,a) && := \lam\,(\gamma[\p]\cdot\q,a[\p]\cdot\q) \\
    & \p && := \lam\,(\fst\,\q) \\
    & \q && := \lam\,(\snd\,\q)

@@

  example we derive the functor law for type substitution as follows.

@@

@@

    & F : \Con_{M'}\ra\Con_M \\
    & F : \Sub_{M'}\,\Delta\,\Gamma\ra\Sub_M\,(F\,\Delta)\,(F\,\Gamma) \\
    & F : \Ty_{M'}\,\Gamma \cong \Ty_M\,(F\,\Gamma) \\
    & F : \Tm_{M'}\,\Gamma\,A \cong \Tm_M\,(F\,\Gamma)\,(F\,A)

@@

  We define them in the same order as follows, omitting the $M$ subscripts.

@@

    & F : \Ty\,\diamond\ra\Con \\
    & F\,\Gamma := \diamond\ext\Gamma \\
    & F : \Tm\,\diamond\,(\Delta\Ra\Gamma) \ra \Sub\,(\diamond\ext\Delta) (\diamond\ext\Gamma) \\
    & F\,\gamma := (\p,\gamma[\p]\cdot\q) \\
    & F : \Tm\,\diamond\,(\Gamma\Ra\U) \cong \Tm\,(\diamond\ext\Gamma)\,\U \cong \Ty\,(\diamond\ext\Gamma) \\
    & F\,A := \El\,(A[\p]\cdot\q) \\
    & F : \Tm\,\diamond\,(\Pi\,\Gamma\,(\El\,(A[\p]\cdot\q))) \cong \Tm\,(\diamond\ext\Gamma)\,(\El\,(A[\p]\cdot\q)) \\
    & F\,a := a[\p]\cdot\q

>>>>

\begin{definition}[Minimised single substitution calculus with $\Pi$ and $\U$]\label{def:min}
  This section defined the minimised version of Definition
  \ref{def:tt}, relying essentially on the presence of universes and
  $\Pi$ types. For reference, we list all the rules of the minimised
  calculus in Appendix \ref{app:minimisation}.
\end{definition}
\begin{problem}
  The GATs of Definition \ref{def:min} and Definition \ref{def:tt} are
  isomorphic, in particular, all equations are interderivable.
\end{problem}
\begin{proof}[Construction]
  Clear from the construction in this section.
\end{proof}

@@

    \Con := (i:\N)\times\Ty\,\diamond\,i \hspace{6em}
    \Sub\,\Delta\,\Gamma := \Tm\,\diamond\,(\Lift^{\Gamma-\Delta}\,\Delta\Ra\Lift^{\Delta-\Gamma}\,\Gamma)

@@

    & \gamma\circ\delta := \lam\Bigg(\mk^{\Theta-\Gamma}\,\bigg(\un^{\Delta-\Gamma}\,\Big(\gamma[\p]\cdot\mk^{\Gamma-\Delta}\,\big(\un^{\Theta-\Delta}\,(\delta[\p]\cdot\mk^{\Delta-\Theta}\,(\un^{\Gamma-\Theta}\,\q))\big)\Big)\bigg)\Bigg) \\
    & \id := \lam\,\q \hspace{10em} \diamond := (0,\top) \hspace{10em} \epsilon := \lam\,(\mk^\Gamma\,\tt)

@@

    & \Tm\,\Gamma\,A && := \Tm\,\diamond\,\Bigg(\Pi\,(\Lift^{i-\Gamma}\,\Gamma)\,\bigg(\Lift^{\Gamma-i}\,\Big(\El\,\big(\un^{\Gamma-(1+i)}\,(A[\p]\cdot\mk^{1+i-\Gamma}\,(\un^{i-\Gamma}\,\q))\big)\Big)\bigg)\Bigg)

@@

    & \Con && := \Ty\,\diamond                                           && \Ty\,\Gamma && := \Tm\,\diamond\,(\Gamma\Ra\U) && \p && := \lam\,(\fst\,\q) \\                     
    & \Sub\,\Delta\,\Gamma && := \Tm\,\diamond\,(\Delta\Ra\Gamma)        && A[\gamma] && := \lam\,(A[\p]\cdot(\gamma[\p]\cdot\q)) && \q && := \lam\,(\snd\,\q) \\
    & \gamma\circ\delta && := \lam\,(\gamma[\p]\cdot(\delta[\p]\cdot\q)) \hspace{1.5em} && \Tm\,\Gamma\,A && := \Tm\,\diamond\,(\Pi\,\Gamma\,(\El\,(A[\p]\cdot\q)))\hspace{1.5em} \\   
    & \id && := \lam\,\q                                                 && a[\gamma] && := \lam\,(a[\p]\cdot(\gamma[\p]\cdot\q)) \\                      
    & \diamond && := \top                                                && \Gamma\ext A && := \Sigma\,\Gamma\,(\El\,(A[\p]\cdot\q)) \\                   
    & \epsilon && := \lam\,\tt                                           && (\gamma,a) && := \lam\,(\gamma[\p]\cdot\q,a[\p]\cdot\q)

@@

  example we derive the functor law for type substitution:

@@

  $

@@

    & F : \Con_{M'}\ra\Con_M && F : \Ty_{M'}\,\Gamma \cong \Ty_M\,(F\,\Gamma) \\
    & F : \Sub_{M'}\,\Delta\,\Gamma\ra\Sub_M\,(F\,\Delta)\,(F\,\Gamma)\hspace{3em} && F : \Tm_{M'}\,\Gamma\,A \cong \Tm_M\,(F\,\Gamma)\,(F\,A)

@@

  We define them in the same order, omitting the $M$ subscripts:

@@

%    & F : \Ty\,\diamond\ra\Con && F : \Tm\,\diamond\,(\Gamma\Ra\U) \cong \Tm\,(\diamond\ext\Gamma)\,\U \cong \Ty\,(\diamond\ext\Gamma) \\
    & F\,\Gamma := \diamond\ext\Gamma && F\,A := \El\,(A[\p]\cdot\q) \\
%    & F : \Tm\,\diamond\,(\Delta\Ra\Gamma) \ra \Sub\,(\diamond\ext\Delta) (\diamond\ext\Gamma) && F : \Tm\,\diamond\,(\Pi\,\Gamma\,(\El\,(A[\p]\cdot\q))) \cong \Tm\,(\diamond\ext\Gamma)\,(\El\,(A[\p]\cdot\q)) \\
    & F\,\gamma := (\p,\gamma[\p]\cdot\q) \hspace{3em} && F\,a := a[\p]\cdot\q
